---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set (
    collapse = TRUE,
    comment = "#>"
)
```

# longtail

Code to analyse GitHub repository activity in relation to "popularity"
metrics.

```{r setup}
devtools::load_all ()
```


## Input data

The chunks below build tables of sample repositories from PyPI and npm,
and full repo details from JOSS and rOpenSci. All data are dumped to the
`OUT_DIR` specified below. These contain repository URLs and popularity
metrics. The section after this then analyses each of the URLs to extract
data on GitHub issue activity.

For PyPI and npm, samples are stratified by download
popularity (a deterministic head of the most-downloaded packages, plus a random
draw from the long tail), filtered down to packages with resolvable GitHub repo
URLs.

## Config

```{r config}
WORKING_SAMPLE_TAIL_SIZE <- 40000L # random draw size, outside the known head
TOP_N_HEAD <- 15000L # deterministic head inclusion
OUT_DIR <- "repo-data-out"
dir.create (OUT_DIR, showWarnings = FALSE)
```

## PyPI

```{r pypi, eval = FALSE}
cli::cli_alert_info ("PyPI: fetching full download-count population via ClickHouse (fast)...")
downloads_tbl <- pypi_downloads_full ()

working_sample <- build_working_sample (downloads_tbl, TOP_N_HEAD, WORKING_SAMPLE_TAIL_SIZE, label = "PyPI")

cli::cli_alert_info ("PyPI: resolving GitHub repo URLs for {nrow(working_sample)} packages...")
pypi_tbl <- resolve_repo_urls (working_sample, pypi_repo_urls_many)
readr::write_csv (pypi_tbl, file.path (OUT_DIR, "pypi.csv"))
cli::cli_alert_success ("PyPI: wrote {nrow(pypi_tbl)} rows to {file.path(OUT_DIR, 'pypi.csv')}")
```

## npm

```{r npm, eval = FALSE}
cli::cli_alert_info ("npm: fetching full download-count population via download-counts package (fast)...")
downloads_tbl <- npm_downloads_full ()

working_sample <- build_working_sample (downloads_tbl, TOP_N_HEAD, WORKING_SAMPLE_TAIL_SIZE, label = "npm")

cli::cli_alert_info ("npm: resolving GitHub repo URLs for {nrow(working_sample)} packages...")
npm_tbl <- resolve_repo_urls (working_sample, npm_repo_urls_many)
readr::write_csv (npm_tbl, file.path (OUT_DIR, "npm.csv"))
cli::cli_alert_success ("npm: wrote {nrow(npm_tbl)} rows to {file.path(OUT_DIR, 'npm.csv')}")
```

## JOSS

Accepted JOSS submissions and their repo URLs.

```{r joss, eval = FALSE}
joss_tbl <- build_joss_table ()
readr::write_csv (joss_tbl, file.path (OUT_DIR, "joss.csv"))
cli::cli_alert_success ("JOSS: wrote {nrow(joss_tbl)} rows to {file.path(OUT_DIR, 'joss.csv')}")
```

## rOpenSci

Packages in the rOpenSci r-universe, their repo URLs, and rOpenSci software
review status.

```{r ropensci, eval = FALSE}
ropensci_tbl <- build_ropensci_table ()
readr::write_csv (ropensci_tbl, file.path (OUT_DIR, "ropensci.csv"))
cli::cli_alert_success ("rOpenSci: wrote {nrow(ropensci_tbl)} rows to {file.path(OUT_DIR, 'ropensci.csv')}")
```
