---
title: "Generating the Data"
author: "Mark Padgham"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Generating the Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set (
    collapse = TRUE, comment = "#>", eval = FALSE
)
```

```{r setup}
library (codingAlone)
```

This vignette walks through the full pipeline used to generate the
`repo-data-out/` dataset that the "review-dividend" vignette analyses:
sampling repositories from PyPI, npm, JOSS, rOpenSci and CRAN, then
fetching every GitHub issue ever opened on each of them. None of the
chunks below are evaluated when this vignette is built - they involve
GitHub API calls that take many hours and require a personal access
token, so this is a reference for reproducing or extending the dataset
yourself, not something that runs automatically.

## Input data

The chunks below build tables of sample repositories from PyPI and npm,
and full repo details from JOSS and rOpenSci. All data are dumped to the
`OUT_DIR` specified below. These contain repository URLs and popularity
metrics. The section after this then analyses each of the URLs to extract
data on GitHub issue activity.

For PyPI and npm, samples are stratified by download
popularity (a deterministic head of the most-downloaded packages, plus a random
draw from the long tail), filtered down to packages with resolvable GitHub repo
URLs.

## Config

```{r config}
WORKING_SAMPLE_TAIL_SIZE <- 40000L # random draw size, outside the known head
TOP_N_HEAD <- 15000L # deterministic head inclusion
OUT_DIR <- "repo-data-out"
dir.create (OUT_DIR, showWarnings = FALSE)
```

## PyPI

```{r pypi}
cli::cli_alert_info ("PyPI: fetching full download-count population via ClickHouse (fast)...")
downloads_tbl <- pypi_downloads_full ()

working_sample <- build_working_sample (downloads_tbl, TOP_N_HEAD, WORKING_SAMPLE_TAIL_SIZE, label = "PyPI")

cli::cli_alert_info ("PyPI: resolving GitHub repo URLs for {nrow(working_sample)} packages...")
pypi_tbl <- resolve_repo_urls (working_sample, pypi_repo_urls_many)
readr::write_csv (pypi_tbl, file.path (OUT_DIR, "pypi.csv"))
cli::cli_alert_success ("PyPI: wrote {nrow(pypi_tbl)} rows to {file.path(OUT_DIR, 'pypi.csv')}")
```

## npm

```{r npm}
cli::cli_alert_info ("npm: fetching full download-count population via download-counts package (fast)...")
downloads_tbl <- npm_downloads_full ()

working_sample <- build_working_sample (downloads_tbl, TOP_N_HEAD, WORKING_SAMPLE_TAIL_SIZE, label = "npm")

cli::cli_alert_info ("npm: resolving GitHub repo URLs for {nrow(working_sample)} packages...")
npm_tbl <- resolve_repo_urls (working_sample, npm_repo_urls_many)
readr::write_csv (npm_tbl, file.path (OUT_DIR, "npm.csv"))
cli::cli_alert_success ("npm: wrote {nrow(npm_tbl)} rows to {file.path(OUT_DIR, 'npm.csv')}")
```

## JOSS

Accepted JOSS submissions and their repo URLs, GitHub stars (fetched via the
GraphQL API), and language label. Also matches repo URLs against the PyPI
and npm tables above (if already written to `OUT_DIR`) to fill in
`downloads` where a JOSS submission's repo happens to also appear in one of
those.

```{r joss}
pypi_csv <- file.path (OUT_DIR, "pypi.csv")
npm_csv <- file.path (OUT_DIR, "npm.csv")
pypi_tbl <- if (file.exists (pypi_csv)) readr::read_csv (pypi_csv, show_col_types = FALSE) else NULL
npm_tbl <- if (file.exists (npm_csv)) readr::read_csv (npm_csv, show_col_types = FALSE) else NULL

joss_tbl <- build_joss_table (pypi_tbl = pypi_tbl, npm_tbl = npm_tbl)
readr::write_csv (joss_tbl, file.path (OUT_DIR, "joss.csv"))
cli::cli_alert_success ("JOSS: wrote {nrow(joss_tbl)} rows to {file.path(OUT_DIR, 'joss.csv')}")
```

## rOpenSci and CRAN

Packages in the any r-universe, their repo URLs, and for rOpenSci, the software
review status.

```{r ropensci}
ropensci_tbl <- build_runiv_table (universe = "ropensci")
readr::write_csv (ropensci_tbl, file.path (OUT_DIR, "ropensci.csv"))
cli::cli_alert_success ("rOpenSci: wrote {nrow(ropensci_tbl)} rows to {file.path(OUT_DIR, 'ropensci.csv')}")

cran_tbl <- build_cran_table ()
readr::write_csv (cran_tbl, file.path (OUT_DIR, "cran.csv"))
cli::cli_alert_success ("CRAN: wrote {nrow(cran_tbl)} rows to {file.path(OUT_DIR, 'cran.csv')}")
```

---

## Analyses

Combine the four input tables into a single table of names, GitHub URLs, and
download/star metrics, tagged with their `source`.

```{r}
repo_tbl <- build_repo_tbl (OUT_DIR)
repo_tbl
```

## GitHub issue authors

Fetch issue-author data for every repo in `repo_tbl` via
`github_issue_authors()`, which returns `repo_url` as one of its columns
directly, along with each repo's own GitHub creation date
(`repo_created_at`, one extra cheap API call per repo) - used by the
issue-rate analysis below as the start of a repo's exposure window.
`repo_url` is the natural unique identifier here (rather than `name`, which
is only unique within a single `source`), so `repo_tbl` is deduplicated on
it before joining, so a repo appearing under multiple sources doesn't fan
out the join.

Calls are made sequentially in small batches. Progress is checkpointed to disk
after every batch, and repos already done are skipped on re-running the chunk,
so an interrupted run - rate-limited or otherwise - just picks back up rather
than starting over.

```{r issue-authors}
issue_authors_tbl <- fetch_issue_authors (repo_tbl$repo_url, OUT_DIR)
issue_authors_tbl <- join_repo_metadata (issue_authors_tbl, repo_tbl)
issue_authors_tbl
```

## Issue-rate analysis

For each fully-fetched source, compute the monthly issue-opening rate
(non-contributor issues per repo-month) stratified by popularity (`downloads`
for `npm`, `stars` for `joss`), fit the interaction model testing whether
that rate trends differently across popularity strata, and plot it.

```{r activity-npm}
rate_tbl_npm <- issue_rate_tbl (issue_authors_tbl, repo_tbl, "npm")
summary (fit_activity_model (rate_tbl_npm))
plot_activity (rate_tbl_npm)
```

```{r activity-joss}
rate_tbl_joss <- issue_rate_tbl (issue_authors_tbl, repo_tbl, "joss")
summary (fit_activity_model (rate_tbl_joss))
plot_activity (rate_tbl_joss)
```
